%%
%% ECE 50874: Solo Project - Failure Analysis
%% Format: ACM single-column (acmart)
%%
\documentclass[manuscript,screen,nonacm]{acmart}

%% ==========================================
%% 1. PACKAGES & CONFIGURATION
%% ==========================================
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{float}
\usepackage{tcolorbox}
\usepackage{cleveref}

\usepackage{xspace}

\newcommand{\ie}{\textit{i.e.}\@\xspace}
\newcommand{\eg}{\textit{e.g.}\@\xspace}

\raggedbottom

\settopmatter{printacmref=false}
\AtBeginDocument{\providecommand\BibTeX{{Bib\TeX}}}

%% ==========================================
%% 2. INSTRUCTION COMMAND
%% ==========================================
\newcommand{\instruction}[1]{%
    \begin{list}{}{%
        \leftmargin=10pt
        \rightmargin=10pt
        \topsep=1em
    }
    \item[]
    \begingroup
        \color{gray}\itshape
        \raggedright
        \sloppy
        \textbf{Instructions:} #1
    \endgroup
    \end{list}
}

\newcommand{\labTwoForumURL}{https://purdue.brightspace.com/d2l/le/1489299/discussions/topics/879296/View}
\newcommand{\labTwoForumLink}{\href{\labTwoForumURL}{Lab 2 Team Signup Discussion (Brightspace)}}


%% ==========================================
%% 3. DOCUMENT METADATA
%% ==========================================
\title{Lab 2: Diverse and Redundant Software Design}
\subtitle{LLM-Powered N-Version Programming}

\begin{document}

%% Author Information
\author{Ming Jia}
\affiliation{%
  \institution{Purdue University}
  \city{West Lafayette}
  \state{Indiana}
  \country{USA}
}
\email{jia213@purdue.edu}

\renewcommand{\shortauthors}{Ming Jia}
%% ==========================================


\begin{abstract}
This is the first lab for the module on Design and Implementation.
The topic is the use of modern large language models (LLMs), or agents based thereupon, to realize the old dream of N-version programming.
See IEC 61508, Part 3, Table A.2, row 3d.

This lab has three parts:
  pre-lab,
  in-class component,
  and take-home component.
All students enrolled in the course are expected to complete the pre-lab and the in-class component.
The take-home component is one of your options for the ``complete 3 take-home labs'' aspect of the course.
\end{abstract}

\maketitle

\section{Lab Purpose and Context}

Earlier in this module, we examined IEC~61508’s treatment of redundancy and diversity in hardware and software. 
We saw that:

\begin{itemize}
  \item Hardware redundancy is often justified using probabilistic independence assumptions.
  \item Software design diversity (IEC~61508 Part~3, Table~A.2, row~3d) --- typically known as \textbf{$N$-version programming} --- is recommended more cautiously.
\end{itemize}

We also reviewed the mathematical argument behind redundancy, which hinges on independence.
Let $F_i$ denote the event that version $i$ fails on a particular input.
\textbf{Independence} means that for all events $i \neq j$, the occurrence of one failure event does not imply the occurrence of another, \ie:

\[
P(F_i \cap F_j) = P(F_i) P(F_j)
\]

If two channels each fail with probability $p$, and failures are independent, then the joint failure probability is $p^2$.
Generalizing this model, voting architectures constructed out of components with independent failure rates (\eg 1-out-of-2/1oo2, 2-out-of-3/2oo3) dramatically reduce the probability of failure.

However, as discussed in class, this independence assumption is questionable in software.
Knight and Leveson (1990) spearheaded the experimental evaluation of $N$-version programming and found substantial correlation in failures across independently-developed programs based on the same specification.
Their results challenged the idea that independent software development naturally produces independent software faults.

In this lab, you will run a modern version of that experiment.

\subsection{From Knight \& Leveson to LLMs}

Knight and Leveson’s original experiment required human teams independently implementing the same specification.
Such arrangements are expensive and have not been found to offer the necessary failure independence to obtain the polynomial reduction in failure when combined.
However, in the modern era, large language models (LLMs) allow us to generate multiple implementations of the same specification at very low marginal cost.
This raises a natural question:

\begin{quote}
\textit{Can LLMs cheaply realize the old dream of $N$-version programming?}
\end{quote}

More precisely, do independently generated implementations by LLMs exhibit the failure independence required for voting-based reliability improvements?

\subsection{Experimental Questions}

This lab investigates two claims:

\begin{enumerate}
  \item \textbf{Independence Claim.} 
  Do LLM-generated implementations exhibit statistically independent failures, as assumed in classical $N$-version programming?

  \item \textbf{Reliability Improvement Claim.}
  Even if failures are correlated, as Knight \& Leveson found, does $N$-version voting improve observed reliability relative to a single implementation?
\end{enumerate}

\subsection{System Under Study}

You will be given a frozen specification for a program.

You may not modify the specification.

Using an LLM or agent system of your choice, you will generate $N$ distinct implementations of the same specification.
Each implementation must satisfy the same functional requirements, but may be produced using different prompting strategies or generation workflows.
This follows the definition of \emph{design diversity}: two or more items carrying out the same function but differing in design  

\subsection{Development and Evaluation Structure}

To mirror the structure of Knight \& Leveson’s experiment:

\begin{itemize}
  \item You will have access to a \textbf{visible test suite} during development.
  \item You may iteratively refine prompts and LLM/agent configurations using feedback from this suite.
  \item A \textbf{reserved test set} will be held back and used only for final evaluation.
\end{itemize}

The specification remains fixed throughout.
The reserved inputs will be used to measure:
\begin{itemize}
  \item Individual failure rates,
  \item Joint failure rates across versions,
  \item Observed reliability of $k$-out-of-$N$ voting.
\end{itemize}


In this lab, you will empirically estimate:

\begin{itemize}
  \item Marginal failure probability,
  \item Joint failure probability,
  \item Conditional failure probability.
\end{itemize}

You will compare observed joint failure rates to those predicted under independence.
This directly operationalizes the mathematical reasoning behind redundant computation from the lecture.

\subsection{Learning Objectives}

After completing this lab, you should be able to:

\begin{itemize}
  \item Explain the mathematical basis of redundancy under independence.
  \item Critically evaluate the independence assumption in software design diversity.
  \item Design and conduct an empirical study evaluating correlated failures.
  \item Use LLMs for simple tasks.
  \item Analyze the effectiveness of $N$-version voting in the presence of correlation.
  \item Connect empirical results to IEC~61508’s guidance on design diversity.
\end{itemize}


\input{1-pre-lab}
\input{2-in-class}
\input{3-take-home}
\clearpage
\input{appendix}

%% ==========================================
%% BIBLIOGRAPHY
%% ==========================================
\clearpage
\nocite{source_example}
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs} 

\end{document}